Data exploration, quality control and statistical analyses of ChIP-exo experiments\vspace*{\fill}

### 2015-11-26 rmarkdown/prelims_paper.Rnw
Changed section and subsection titles

\title{High Resolution Identification of Protein-DNA Binding Events
  and Quality Control for ChIP-exo data\vspace*{\fill}}

@@

Currently, research on statistical methods for ChIP-exo data is still
in its very early stage. Although many methods have been proposed to
identify protein binding sites from ChIP-Seq data (reviewed by
Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
2009 \cite{computation}), such as MACS (Zhang et al., 2008
\cite{macs}), CisGenome (Hongkai et al., 2008 \cite{cisgenome}) and
MOSAiCS (Kuan et al., 2009 \cite{mosaics}), these approaches reveal
protein binding sites in lower resolution, i.e., at an interval of
hundreds to thousands of base pairs. Furthermore, they report only one
``mode'' or ``predicted binding location'' per peak. Hence, these
methods are not appropriate to evaluate the potential of ChIP-exo data
for high resolution identification of protein binding sites. More
recently, deconvolution algorithms such as Deconvolution (Lun et al.,
2009 \cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
\cite{pics}) have been proposed to identify binding sites in higher
resolution using ChIP-Seq data. However, most of them are still not
tailored for ChIP-exo and PET and SET ChIP-Seq data in a unified
framework and as a result, currently available methods are not
appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
address these limitations, we developed an improved of dPeak (Chung et
al., 2013 \cite{dpeak}), a high resolution binding site identification
(deconvolution) algorithm that we previously developed for PET and SET
ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
algorithm implements a probabilistic model that accurately describes
the ChIP-exo and ChIP-Seq data generation process.

@@

\subsubsection{Strand cross-correlation}
\label{sec:scc}

@@

\subsection{The dPeak model}
\label{sec:dpeak}

@@

\subsubsection{dPeak's algorithm initialization strategy}
\label{sec:algo}

@@

\subsection{ChIP-exo Quality Control pipeline}
\label{sec:QC}

@@

\subsubsection{Local Normalized Strand cross-correlation}
\label{sec:locNSC}

@@

\subsubsection{Enrichment and library complexity in ChIP-exo data}
\label{sec:enri}

@@

\subsubsection{Strand imbalance in ChIP-exo data}
\label{sec:strand_imbalance}

@@

\subsection{Recommendations for the design of ChIP-exo experiments}
\label{sec:reco}

@@

\subsection{Some additional thoughts for this analysis}

\subsubsection{GC content and mappability quality indicators}
\label{sec:gc_map_qc}

@@

\subsubsection{Quality control for ChIP-Nexus}
\label{sec:nexus}

@@

\subsection{Analysis of E. Coli Transcription Initiation Complexes}
\label{sec:ecoli}

@@

\subsection{Enhancer prediction and active learning}
\label{sec:enhancer}

>>>>

\title{Data exploration, quality control and statistical analyses of
  ChIP-exo experiments\vspace*{\fill}}

@@

Currently, research on statistical methods for ChIP-exo data is still
in its very early stage. Although many methods have been proposed to
identify protein binding sites from ChIP-Seq data (reviewed by
Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
2009 \cite{computation}), such as MACS (Zhang et al., 2008
\cite{macs}), CisGenome (Ji et al., 2008 \cite{cisgenome}) and
MOSAiCS (Kuan et al., 2009 \cite{mosaics}), these approaches reveal
protein binding sites in lower resolution, i.e., at an interval of
hundreds to thousands of base pairs. Furthermore, they report only one
``mode'' or ``predicted binding location'' per peak. Hence, these
methods are not appropriate to evaluate the potential of ChIP-exo data
for high resolution identification of protein binding sites. More
recently, deconvolution algorithms such as Deconvolution (Lun et al.,
2009 \cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
\cite{pics}) have been proposed to identify binding sites in higher
resolution using ChIP-Seq data. However, most of them are still not
tailored for ChIP-exo and PET and SET ChIP-Seq data in a unified
framework and as a result, currently available methods are not
appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
address these limitations, we developed an improved of dPeak (Chung et
al., 2013 \cite{dpeak}), a high resolution binding site identification
(deconvolution) algorithm that we previously developed for PET and SET
ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
algorithm implements a probabilistic model that accurately describes
the ChIP-exo and ChIP-Seq data generation process.

@@

\subsubsection{Strand Cross-Correlation}
\label{sec:scc}

@@

\subsection{The dPeak Model}
\label{sec:dpeak}

@@

\subsubsection{A New Initialization Strategy for dPeak}
\label{sec:algo}

@@

\subsection{ChIP-exo Quality Control Pipeline}
\label{sec:QC}

@@

\subsubsection{Normalized Local Strand Cross-Correlation}
\label{sec:locNSC}

@@

\subsubsection{Enrichment and Library Complexity in ChIP-exo data}
\label{sec:enri}

@@

\subsubsection{Strand Imbalance in ChIP-exo data}
\label{sec:strand_imbalance}

@@

\subsection{Systematic Comparison of ChIP-Seq vs ChIP-exo under
  Varying Sequencing Depths}
\label{sec:reco}

@@

\subsection{Additional Thoughts for this Analysis}

\subsubsection{GC Content and Mappability Quality Indicators}
\label{sec:gc_map_qc}

@@

\subsubsection{Quality Control for ChIP-Nexus}
\label{sec:nexus}

@@

\subsection{Analysis of \emph{E. Coli} Transcription Initiation Complexes}
\label{sec:ecoli}

@@

\subsection{Enhancer Prediction and Active Learning}
\label{sec:enhancer}
